Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-08-15 Manuscript/manuscript.tex
Corrected surface excess Equation and statements around it. Surface excess was not written correctly, factor 1/2 missing in the final form for simulations, concentrations use same symbols as in previous paragraph. Corrected reference for chattorajBOOK.

The experimental data used to quantify cation binding reports
order parameters as a function of equilibrium cation concentration
in the bulk solvent \cite{akutsu81,altenbach84}. Recent work
by NMRlipids project estimated this concentration by using
the relative molar concentration in water
[salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M,
which was accurate enough for the conclusions in that study \cite{catte16}.
Since we are now targeting a model with quantitatively accurate binding
affinity, we rather use a definition that corresponds more closely to the 
concentrations reported in experiments. 
By using larger simulation box dimensions that allow the
determination of bulk ion conceration, the equilibrium concentration
corresponding to the experimental data in the units of $\mathrm{mol/l}$
can be then determined from the simulations as
\begin{equation}
\label{eq:conc}
  C_{eq}=\frac{C_{bulk}}{0.602},
\end{equation}
where bulk concentration, $C_{bulk}$, is the number density in the units of $\mathrm{nm}^{-3}$.
\todo{SAMULI: Once we have to final results, we can probably say that
  the repeat distance is not far from the experimentally measured distance \cite{pabst07,uhrikova08}
  }

The ion binding affinities between different models were compared by calculating
the relative surface excess between water and ions $\Gamma_w^i$ \cite{chattorajBOOK}
\begin{equation}
  \Gamma_w^i=\Gamma_w^x-\Gamma_i^x\frac{c_w-c_w'}{c_i-c_i'},
\end{equation}
where $c_w$ and $c_i$ are concentrations in bulk water, 
$c_w'$ and $c_i'$ are concentrations inside bilayer, and
$\Gamma_w^x$ and $\Gamma_i^x$ are surface excess of water and
ions, respectively. The whole simulation box is considered here
as interfacial region, thus $\Gamma_w^x=n_w/A$ and $\Gamma_i^x=n_i/A$,
where $A$ is the area of the simulation box, and $n_w$ and
$n_i$ are total number of water and ions in simulation box,
respectively. The concentrations of water and ions inside
bilayer can be expected to vanish, thus $c_w'=c_i'=0$.
Then we get the relative surface excess for lipid bilayer
simulations
\begin{equation}\label{surfexcess}
  \Gamma_w^i=\frac{n_w}{A}-\frac{n_i}{A}\frac{c_w}{c_i},
\end{equation}
where concentrations in bulk, $c_w$ and $c_i$, can be determined
from the simulation box boundary.

>>>>

The experimental data used to quantify cation binding reports
order parameters as a function of equilibrium cation concentration
in the bulk solvent \cite{akutsu81,altenbach84}. Recent work
by NMRlipids project estimated this concentration by using
the relative molar concentration in water
[salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M,
which was accurate enough for the conclusions in that study \cite{catte16}.
Since we are now targeting a model with quantitatively accurate binding
affinity, we rather use a definition that corresponds more closely to the 
concentrations reported in experiments. 
By using larger simulation box dimensions that allow the
determination of bulk ion concentration, the equilibrium concentration
corresponding to the experimental data in the units of $\mathrm{mol/l}$
can be then determined from the simulations as
\begin{equation}
\label{eq:conc}
  C_{eq}=\frac{C_{bulk}}{0.602},
\end{equation}
where the bulk concentration, $C_{bulk}$, is the number density in the units of $\mathrm{nm}^{-3}$
in the bulk solvent far from the membrane interface. 
\todo{SAMULI: Once we have to final results, we can probably say that
  the repeat distance is not far from the experimentally measured distance \cite{pabst07,uhrikova08}
  }

The ion binding affinities between different models were compared by calculating
the relative surface excess between water and ions $\Gamma_w^i$ \cite{chattorajBOOK}
\begin{eqnarray}
  \Gamma_w^i   &=& \Gamma_w-\Gamma_i \frac{C_w-C_w'}{C_i-C_i'} \\
  \Gamma_{a} &=& \frac{n_{a} - C_{a} V - C_{a}' V'}{A} \, \, ; \, a=i,w \, ,
\end{eqnarray}
where index $i$ and $w$ denote ions and water respectively,
prime accent denotes the phase inside bilayer, 
whereas non-prime symbols account for bulk water phase,
$n$ is the total amount of water or ions,
$C$ denotes bulk concentrations, 
$V$ is the volume of the respective phase separated by the Gibbs divinding plane, 
$A$ is the area of the interface, 
and $\Gamma_a$ is the surface excess of water or ions. 
As $\Gamma_w^i$ does not depend on the definition of the interfacial region
and the position of the Gibbs dividing plane~\cite{chattorajBOOK},
we consider the whole simulation box as an interface. 
Note that phospholipid membranes are bilayers, 
and hence the interface occurs in the simulation box twice
providing a surface area $2A$.
The concentrations of water and ions inside
the bilayer can be expected to vanish, thus $c_{i,w}'=0$.
This provides us a simplified relation for 
the relative surface excess of lipid bilayers in simulations
\begin{equation}\label{surfexcess}
  \Gamma_w^i=\frac{1}{2A} \left ( n_w - n_i \frac{C_w}{C_i} \right ) ,
\end{equation}
where the bulk concentrations, $C_w$ and $C_i$, can be determined
from the simulation as $C_{bulk}$ in Equation~\ref{eq:conc}.
